Incremental Faint Object Detection in Multi-Epoch Observations

### 2016-11-04 ms.tex
Many small clean-ups; new refs

\begin{abstract} 
Observational astronomy in the time-domain era faces several new challenges. 
One of them is the efficient use of observations obtained at multiple epochs. 
The work presented here addresses faint object detection, and seeks to find an incremental strategy for separating real objects from artifacts in ongoing surveys, in situations where the available data are summaries of the full image data, such as source flux and direction estimates (with uncertainties) reported in catalogs.
We adopt a Bayesian approach, addressing object detection by calculating the probabilities for hypotheses asserting there is no object, or one object, in a small image patch containing at most one detected source at each epoch.
The object-present hypothesis interprets the sources in a patch as arising from a genuine object; the no-object hypothesis interprets any detected sources as spurious, arising from noise peaks.
We study the detection probability for constant-flux objects in a simplified Gaussian noise setting, comparing results based on single exposures and stacked exposures to results based on catalog summary data.
Computing the detection probability based on catalog data amounts to generalized cross-matching: it is the product of a factor accounting for matching of the estimated fluxes of candidate sources, and a factor accounting for matching of their estimated directions (i.e., directional cross-matching across catalogs).
We find that probabilistic fusion of multi-epoch catalog information can detect sources with only modest sacrifice in sensitivity compared to stacking.
We show that the probabilistic cross-matching underlying our approach plays an important role in maintaining detection sensitivity.
\end{abstract}

@@

There are two different ways one can approach the problem. 
A traditional, resource-intensive approach is to wait until all observations are completed, performing detection by stacking the multi-epoch image data.
An optimal procedure for threshold-based detection with image stacks was introduced by \citet{chisq}.
Once a master object catalog is produced from the stacked images, time-series of source measurements are created by forced photometry at the master catalog locations.

@@

We here address the second alternative, considering how best to use the list of detections while the observations are in progress to prune spurious source detections but keep the sources associated with genuine objects.
The study presented here is exploratory, to establish the basic ideas and provide initial metrics for studying the feasibility of the incremental approach.
To make the analysis analytically tractable and the results easy to interpret, we restrict ourselves to an idealized setting; we will present a more general treatment in a subsequent paper.

@@

We consider a catalog at a given epoch to report summaries of the likelihood functions for candidate sources that have met some detection criteria.
The most commonly reported summaries are best-fit fluxes (or magnitudes) with a quantification of flux uncertainty (typically a standard error or the half-width of a 68\% confidence region), and, separately, best-fit sky coordinates with an uncertainty for each coordinate.%
%***\footnote{Cautionary note about the meaning of ``best-fit'' here.}
We here take these summaries to correspond to a factored approximation of the source likelihood function,
\begin{align} \label{eq:eplike}
\like_\eind(\flux, \drxn)
  &\equiv p(\edata|\flux, \drxn, H_1)\nonumber\\
  &= \flike_\eind(\flux)\, \dlike_\eind(\drxn),
\end{align}
where the epoch-specific flux factor, $\flike_\eind(\flux)$, is a Gaussian with mode $\fest_\eind$ (the catalog flux estimate at epoch $\eind$) and standard deviation $\fsig_\eind$, and the direction factor, $\dlike_\eind(\drxn)$, is an azimuthally symmetric bivariate Gaussian with mode $\dest_\eind$, and standard deviation $\dsig_\eind$.
This may be a rather crude approximation; we will address it further elsewhere, here merely noting that it is implicitly adopted for most survey catalogs.
For simplicity, we take the flux factors to have the same standard deviation at all epochs,  $\fsig_\eind = \sigma$.
\enote{Define the direction uncertainty, too.}

@@

Figure~\ref{fig:3} compares detection probability curves for the stacked exposure case (solid yellow curves, as in Figure~\ref{fig:1}) and the multi-epoch, $n>n_0$ detection case (dashed blue curves, as in Figure~\ref{fig:2}).
For a particular stacked exposure case, we see there is a multi-epoch case whose detection probability curve displays very similar performance.
For example, the $3\sigma$ stacked exposure curve is very similar to the multi-epoch $5+$ detection case.
This indicates that collecting sources with $5+$ detections from \emph{single-epoch} $3\sigma$ catalogs is nearly equivalent in terms of catalog completeness and purity to producing a separate, new $3\sigma$ stacked exposure catalog.
%
\enote{Move the rest to intro?}
Analyzing the single-epoch catalogs has a number of advantages.
It can be implemented in an incremental fashion that follows the schedule of the survey and the time-series is readily available at a given time; there is no need to go back to old images and to performed forced photometry at locations that are revealed in the stacks only.

@@

%================================================================================
\section{Distinguishing Real Sources from Noise Peaks}
\noindent
%
The previous calculations addressed detectability of a source of known true flux, $f$.
In real-life scenarios, the problem is quite the opposite---we are presented with the observations and would like to understand the properties of the sources. 
In this context, our focus is on how one can reliably distinguish noise peaks from real sources. 
It is important to emphasize that we have more information than just the fact that a source has been detected; we also have flux measurements, at multiple epochs. 
Our approach is motivated by Bayesian hypothesis testing, where the strength of evidence for presence of a source is quantified by the posterior probability for the source-present hypothesis, or equivalently, by the posterior odds in favor of a source being present vs.\ being absent (the odds is the ratio of probabilities for the rival hypotheses).
The posterior odds is the product of prior odds and the data-dependent \emph{Bayes factor}.
The prior odds depends on population properties; it may be specified a priori when there is sufficient knowledge of the population under study, or learned adaptively by using hierarchical Bayesian methods (for examples of this in the related context of cross-identification, see REF).
\enote{Get SCMA5 ref.}
Here we focus on the Bayes factor; we will address hierarchical modeling in a follow-up paper.

\enote{Should ``source'' be ``object'' here?}

The Bayes factor is the ratio of marginal likelihoods for the competing hypotheses, one that claims that the source is real, and its complement that assumes just noise:
%
\begin{equation} \label{eq:Bfac}
B = \frac{\mlike_{\rm{}real}}{\mlike_{\rm{}noise}} .
\end{equation}
%
Each marginal likelihood, $\mlike_{\rm{}hyp}$, is the integral, with respect to all free parameters for the hypothesis, of the product of the likelihood function and the prior probability density for the parameters.

@@

%--------------------------------------------------------------------------------
\subsection{Real-source hypothesis}
\noindent
Let $\dtxn$ denote the set of indices for epochs with detections, and $\ndtxn$ denote the set of indices for epochs with nondetections:
\begin{equation}
\begin{split}
\dtxn  &\equiv \{\eind : \fest_\eind > \fth\},\\
\ndtxn &\equiv \{\eind : \fest_\eind \le \fth\}.
\end{split}
\label{D-ND-sets}
\end{equation}
For a candidate source with $n$ detections among $k$ catalogs, the likelihood for a candidate true flux $f$ is
%
\begin{equation}
\like(f) = (1\!-\!P_f)^{k-n}\,\prod_{\eind \in \dtxn} \flike_\eind(\flux),
\end{equation}
%
where \mbox{$(1\!-\!P_f)$} is the probability of not detecting an object with true flux $f$, which happens \mbox{$(k\!-\!n)$} times, and $\flike_\eind(\cdot)$ is the flux likelihood function defined above (Gaussians with means equal to $\fest_\eind$).
%
The marginal likelihood for the real-source hypothesis is obtained by averaging over all possible true flux values, $f$.
For a source that is a member of a population with known flux probability density $\pi(f)$, the prior probability for $f$, used for the averaging in the marginal likelihood, is $\pi(f)$, so that
%
\begin{equation}
\mlike_{\rm{}real} = \int\!\!df\,\pi(f)\,L(f)
\end{equation}
%
which is a one-dimensional integral that can be analytically or numerically quickly evaluated.
(When the population distribution is not known a priori, it may be estimated via joint analysis of the catalog data for many candidate sources, within a hierarchical model, a significant complication that we will elaborate on elsewhere.)

%--------------------------------------------------------------------------------
\subsection{Noise peak hypothesis}\label{sec:peaks}
\noindent
%
The alternative hypothesis is that the detections are simply random noise peaks in the image.
The noise hypothesis marginal likelihood, $\mlike_{\rm noise}$, is the probability for the catalog data presuming no real source is present.\footnote{We are presuming the use of a fixed source detection algorithm, e.g., fixed apertures. 
If the algorithm is adaptive, its tuning parameters need to be accounted for in this probability.}

@@

As noted above, we obtain the probability for detecting a noise peak, $P_N$, by integrating $\npd(\fest)$ above the flux threshold.
%
Figure~\ref{fig:frac}a shows the results as a function of the flux threshold in $\sigma$ units (left), and on a scale corresponding to an LSST-like magnitude (right).
\enote{We should specify how the LSST magnitude scale was set, or point to a REF.}
We see that the fraction of noise peaks above threshold is about 62\% at 1$\sigma$, dropping quickly to about a few percent at 3$\sigma$, and becoming negligible at 5$\sigma$.
%
Based on just this figure, it is tempting to set a high detection threshold to reject such ``ghost peaks'' and keep the catalog of detections nearly pure; but that would mean we lose the opportunity to recover the numerous really faint sources.

@@

The detections from a real source are all connected, they are just displaced by a random astrometric error; but noise peaks (ghosts) will be independent of each other and their associations can only be by chance.
As we are working under the approximation that the flux and sky position estimates are independent (see Eq.~\ref{eq:eplike}), the Bayes factor using both the photometric and astrometric information factors,
%
\begin{equation}
B_{\rm{}flux,pos} = B_{\rm{}flux}\cdot{}B_{\rm{}pos}.
\end{equation}
%
\enote{Find REF.}
The astrometric cross-match Bayes factor, $B_{\rm{}pos}$, has been derived in \ref{XX} (see Eq.~(17) there, and Eq.~(19) for the tangent plane Gaussian limit that holds for high-precision astrometry).
That work also discusses generalizations that account for proper motion and other complications.

@@

\noindent
We here use simulations to demonstrate the detection capability of our multi-epoch approach in a setting with known ground truth.
The simulation parameters were chosen to produce data similar to that provided by modern large-scale optical surveys.

%--------------------------------------------------------------------------------
\subsection{Simulated Galaxies}
\noindent
We assume that galaxies are brighter than 28 magnitudes and that the 5$\sigma$ detection limit is 24 magnitudes, corresponding roughly to parameters of LSST photometry.
%
Panel~(b) of Figure~\ref{fig:frac} shows the noise peak detection probability as a function of magnitude based on these parameters, in contrast to the dimensionless presentation in panel~(a).
%
To compute the marginal likelihood for the source-present hypothesis, we must specify a prior for the source flux, $\flux$.
Here we use a standard faint-galaxy number counts law,
with the number counts following the empirical formula of 
\mbox{$dN\!\propto\!10^{0.4m}dm$}; see \cite{MT00-NumCounts}. 
That approximately translates to the properly normalized population distribution of
%
\begin{equation} 
\label{eq:ref3}
\pi(f) = \left\{\begin{array}{l l}
           f_L/f^2  & \quad \mbox{if\ $f > f_L$}\\
           0 & \quad \mbox{otherwise}
           \\ \end{array} \right.
\end{equation}
%
where $f_L$ is the limiting flux that corresponds to the previously defined magnitude limit.

%\subsection{The Mock Catalogs}\label{sec:mock} \noindent
We generate sets of random detections for 20,000 galaxies with true fluxes between 28 and 23 magnitudes by simply drawing $\fest_\eind$ values from a Gaussian centered on the actual fluxes.
%
We also generate 2,000 ghost detection $\fest_\eind$ values from $\npd(\fest)$ by inverting its cumulative distribution (computed numerically on a grid).
%
The number of exposures is set to the previously used $k=9$ with a flux threshold of just 1$\sigma$, deep in the noise.
In observations with our specified parameters, the number of ghost detections will greatly outnumber the galaxy detections with this low threshold.
The numbers of galaxies and ghosts were chosen to enable display of the distributions of Bayes factors for the two classes of detections (noise and true).

@@

This computation demonstrates that flux and astrometric catalog data, combined across epochs, can strongly distinguish real objects from spurious detections.
We have not addressed what threshold Bayes factor to use for producing a multi-epoch catalog, or, in Bayesian terms, how to convert Bayes factors into posterior probabilities for candidate objects.
As noted above, when the object population density (on the sky and in flux) is known a priori, the calculation is straightforward (e.g., the prior odds will be proportional to the ratio of true object and noise peak sky densities).
When these quantities are unknown, a possibly complicated hierarchical Bayesian calculation can jointly estimate them and the object properties.
When many objects are detected, an approximate approach, plugging in empirical estimates of the densities based on the data, is likely to suffice, as described in \cite{REF,REF}.
\enote{SCMA5 refs?}

%****************************************
\color{red}
\begin{figure}
\epsscale{1.2}
\plotone{fig/f7u-U.png}
\caption{The (red) points represent the weight of evidence for simulated galaxies as a function of their true brightness and the same for generated random noise peaks. The Bayes factor is high for the bright galaxies but its logarithm is close to zero for the faint ones. The noise peaks (left of double lines) seem to separate slightly from the faintest sources (just right of the double dashed vertical lines) because the fluxes in noise peaks tend to be higher, which better mimics sources between 26 and 25 magnitudes.}
\label{fig:bf}
\end{figure}

@@

%================================================================================
\section{Summary}
\label{sec:sum}
\noindent
In this paper we explored new approaches to processing a series of exposures with special focus on the faintest sources, where most of the new discoveries are expected.
%
We found that one can recover sources just by collecting a number of detections above even a low threshold with the same probability as if the sources were extracted from coadded images. Consequently we can develop incremental algorithms that crossmatch the detections and weed out the noise on the fly from a master source catalog.

We derived the spatial properties of noise peaks that commonly appear in catalogs.
The surface density of these ghosts is asymmetric and skewed toward positive flux values but can be accurately approximated by a shifted Gaussian for most practical purposes.
We used our analytic results to perform Bayesian hypothesis testing to distinguish between true sources and noise peaks.
%
Based on the Bayes factor, the bright sources over \mbox{3$\sigma$} \mbox{($m\!\simeq\!24.55$)} start to separate out from the noise peaks, which becomes with very convincing evidence at \mbox{5$\sigma$} (24 mags).
%
Just based on the flux measurements, the faint sources are practically indistinguishable from the noise peaks but their celestial coordinates can help to correctly identify them as real sources. Using probabilistic cross-identification is to our direct benefit directly providing the required quantities.
For best results we combine the brightness and spatial information in a probabilistic manner by multiplying the Bayes factors.

In general, the specificity and the selectivity of the proposed discriminator depends on a number of parameters most of which we discussed as part of the simulated case study.
The pixel size, the detection limit, the number of exposures and the point-spread function will determine the frequency of noise peaks as well as the statistics of the real sources based on which one can come up with reliable thresholds for progressively weeding out noise from the catalogs throughout the lifetime of a survey.

@@

\bibitem[Gregory \& Loredo(1992)]{gregory} Gregory, P.~C., \& Loredo, T.~J.\ 1992, \apj, 398, 146

@@

\bibitem[Loredo(2012)]{loredo} Loredo, T.~J.\ 2012, arXiv:1206.4278

@@

\appendix
\color{blue}

>>>>

\begin{abstract} 
Observational astronomy in the time-domain era faces several new challenges. 
One of them is the efficient use of observations obtained at multiple epochs. 
The work presented here addresses faint object detection with multi-epoch data, and describes an incremental strategy for separating real objects from artifacts in ongoing surveys, in situations where the single-epoch data are summaries of the full image data, such as single-epoch catalogs of flux and direction estimates (with uncertainties) for candidate sources.
The basic idea is to produce low-threshold single-epoch catalogs, and use a probabilistic approach to accumulate catalog information across epochs; this is in contrast to more conventional strategies based on co-added or stacked image data across all epochs.
We adopt a Bayesian approach, addressing object detection by calculating the marginal likelihoods for hypotheses asserting there is no object, or one object, in a small image patch containing at most one cataloged source at each epoch.
The object-present hypothesis interprets the sources in a patch at different epochs as arising from a genuine object; the no-object (noise) hypothesis interprets candidate sources as spurious, arising from noise peaks.
We study the detection probability for constant-flux objects in a simplified Gaussian noise setting, comparing results based on single exposures and stacked exposures to results based on a series of single-epoch catalog summaries.
Computing the detection probability based on catalog data amounts to generalized cross-matching: it is the product of a factor accounting for matching of the estimated fluxes of candidate sources, and a factor accounting for matching of their estimated directions (i.e., directional cross-matching across catalogs).
We find that probabilistic fusion of multi-epoch catalog information can detect sources with only modest sacrifice in sensitivity compared to stacking.
The probabilistic cross-matching framework underlying our approach plays an important role in maintaining detection sensitivity, and points toward generalizations that could accomodate variability and complex object structure.
\end{abstract}

@@

There are two different ways one can approach the problem. 
A traditional, resource-intensive approach is to wait until all observations are completed, performing detection by stacking the multi-epoch image data (with potential complications related to registration, resampling, and point spread function matching).
An optimal procedure for threshold-based detection with image stacks was introduced by \citet{chisq}.
Once a master object catalog is produced from the stacked images, time-series of source measurements are created by forced photometry at the master catalog locations.

@@

\enote{Maybe clarify here whether we envision this as useful to end-users, or only for survey data management teams. To me it seems we should leave open the former possibility; if so, we may not need to add anything here.}

We here address the second alternative, considering how best to accumulate evidence from possibly marginal detections while the observations are in progress, to prune spurious source detections but keep the sources associated with genuine objects.
The study presented here is exploratory, to establish the basic ideas and provide initial metrics for studying the feasibility of the incremental approach.
To make the analysis analytically tractable and the results easy to interpret, we restrict ourselves to an idealized setting; we will present a more general and formal treatment in a subsequent paper.

@@

We consider a catalog at a given epoch to report summaries of the likelihood functions for candidate sources that have met some detection criteria.
The most commonly reported summaries are best-fit fluxes (or magnitudes) with a quantification of flux uncertainty (typically a standard error or the half-width of a 68\% confidence region), and, separately, best-fit sky coordinates with an uncertainty for each coordinate.%
%***\footnote{Cautionary note about the meaning of ``best-fit'' here.}
We here take these summaries to correspond to a factored approximation of the source likelihood function,
\begin{align} \label{eq:eplike}
\like_\eind(\flux, \drxn)
  &\equiv p(\edata|\flux, \drxn, H_1)\nonumber\\
  &= \flike_\eind(\flux)\, \dlike_\eind(\drxn),
\end{align}
where the epoch-specific flux factor, $\flike_\eind(\flux)$, is a Gaussian with mode $\fest_\eind$ (the catalog flux estimate at epoch $\eind$) and standard deviation $\fsig_\eind$, and the direction factor, $\dlike_\eind(\drxn)$, is an azimuthally symmetric bivariate Gaussian with mode $\dest_\eind$, and standard deviation $\dsig_\eind$.%
\footnote{Care should be taken in reporting and interpreting $\dsig_\eind$, because direction is a two-dimensional quantity.
If $\dsig_\eind$ is the single-coordinate standard deviation, the angular radius of a 68.3\% (``$1\sigma$'') confidence region or flat-prior credible region is $\approx 1.52\dsig_\eind$.}
This may be a rather crude approximation; we will address it further elsewhere, here merely noting that it is implicitly adopted for most survey catalogs.
For simplicity, we take the flux factors to have the same standard deviation at all epochs,  $\fsig_\eind = \sigma$.

@@

Figure~\ref{fig:3} compares detection probability curves for the stacked exposure case (solid yellow curves, as in Figure~\ref{fig:1}) and the multi-epoch, $n>n_0$ detection case (dashed blue curves, as in Figure~\ref{fig:2}).
For a particular stacked exposure case, we see there is a multi-epoch case whose detection probability curve displays very similar performance.
For example, the $3\sigma$ stacked exposure curve is very similar to the multi-epoch $5+$ detection case.
This indicates that collecting sources with $5+$ detections from \emph{single-epoch} $3\sigma$ catalogs is nearly equivalent in terms of catalog completeness and purity to producing a separate, new $3\sigma$ stacked exposure catalog.
%
\enote{Move the rest to intro? It may already be addressed there.}
Analyzing the single-epoch catalogs has a number of advantages.
It can be implemented in an incremental fashion that follows the schedule of the survey and the time-series is readily available at a given time; there is no need to go back to old images and to performed forced photometry at locations that are revealed in the stacks only.

@@

%================================================================================
\section{Distinguishing Real Sources from Noise Peaks}
\noindent
%
The previous calculations addressed detectability of a source of known true flux, $f$.
In real-life scenarios, the problem is quite the opposite---we are presented with the observations and would like to understand the properties of the sources. 
In this context, our focus is on how one can reliably distinguish noise peaks from real sources. 
It is important to emphasize that we have more information than just the fact that a source has been detected; we also have flux measurements, at multiple epochs. 
Our approach is motivated by Bayesian hypothesis testing, where the strength of evidence for presence of a source is quantified by the posterior probability for the source-present hypothesis, or equivalently, by the posterior odds in favor of a source being present vs.\ being absent (the odds is the ratio of probabilities for the rival hypotheses).
The posterior odds is the product of prior odds and the data-dependent \emph{Bayes factor}.
The prior odds depends on population properties; it may be specified a priori when there is sufficient knowledge of the population under study, or learned adaptively by using hierarchical Bayesian methods (for examples of this in the related context of cross-identification, see \cite{B12-SCMAXMatch,L12-HierXMatch}).
Here we focus on the Bayes factor; we will address hierarchical modeling in a follow-up paper.

The Bayes factor is the ratio of marginal likelihoods for the competing hypotheses, one that claims that the sources are associated with a real object, and its complement that assumes there is just noise:
%
\begin{equation} \label{eq:Bfac}
B = \frac{\mlike_{\rm{}real}}{\mlike_{\rm{}noise}} .
\end{equation}
%
Each marginal likelihood, $\mlike_{\rm{}hyp}$, is the integral, with respect to all free parameters for the hypothesis, of the product of the likelihood function and the prior probability density for the parameters.

@@

%--------------------------------------------------------------------------------
\subsection{Real-object hypothesis}
\noindent
Let $\dtxn$ denote the set of indices for epochs with detections, and $\ndtxn$ denote the set of indices for epochs with nondetections:
\begin{equation}
\begin{split}
\dtxn  &\equiv \{\eind : \fest_\eind > \fth\},\\
\ndtxn &\equiv \{\eind : \fest_\eind \le \fth\}.
\end{split}
\label{D-ND-sets}
\end{equation}
For a candidate object with $n$ source detections among $k$ catalogs, the likelihood for a candidate true flux $f$ is
%
\begin{equation}
\like(f) = (1\!-\!P_f)^{k-n}\,\prod_{\eind \in \dtxn} \flike_\eind(\flux),
\end{equation}
%
where \mbox{$(1\!-\!P_f)$} is the probability of not detecting an object with true flux $f$, which happens \mbox{$(k\!-\!n)$} times, and $\flike_\eind(\cdot)$ is the flux likelihood function defined above (Gaussians with means equal to $\fest_\eind$).
%
The marginal likelihood for the real-object hypothesis is obtained by averaging over all possible true flux values, $f$.
For an object that is a member of a population with known flux probability density $\pi(f)$, the prior probability for $f$, used for the averaging in the marginal likelihood, is $\pi(f)$, so that
%
\begin{equation}
\mlike_{\rm{}real} = \int\!\!df\,\pi(f)\,L(f)
\end{equation}
%
which is a one-dimensional integral that can be analytically or numerically quickly evaluated.
(When the population distribution is not known a priori, it may be estimated via joint analysis of the catalog data for many candidate objects, within a hierarchical model, a significant complication that we will elaborate on elsewhere.)

%--------------------------------------------------------------------------------
\subsection{Noise peak hypothesis}\label{sec:peaks}
\noindent
%
The alternative hypothesis is that the detections are simply random noise peaks in the image.
The noise hypothesis marginal likelihood, $\mlike_{\rm noise}$, is the probability for the catalog data presuming no real object is present.%
\footnote{We are presuming the use of a fixed source detection algorithm, e.g., fixed apertures. 
If the algorithm is adaptive, its tuning parameters need to be accounted for in this probability.}

@@

As noted above, we obtain the probability for detecting a noise peak, $P_N$, by integrating $\npd(\fest)$ above the flux threshold.
%
Figure~\ref{fig:frac}a shows the results as a function of the flux threshold in $\sigma$ units (left), and on a scale corresponding to an LSST-like magnitude (right; see \S~\ref{sec:disc} for a description of the magnitude scale).
We see that the fraction of noise peaks above threshold is about 62\% at 1$\sigma$, dropping quickly to about a few percent at 3$\sigma$, and becoming negligible at 5$\sigma$.
%
Based on just this figure, it is tempting to set a high detection threshold to reject such ``ghost peaks'' and keep the catalog of detections nearly pure; but that would mean we lose the opportunity to recover the numerous really faint sources.

@@

The detections from a real source are all connected, they are just displaced by a random astrometric error; but noise peaks (ghosts) will be independent of each other and their associations can only be by chance.
As we are working under the approximation that the flux and sky position estimates are independent (see Eq.~\ref{eq:eplike}), the Bayes factor using both the photometric and astrometric information factors,
%
\begin{equation}
B_{\rm{}flux,pos} = B_{\rm{}flux}\cdot{}B_{\rm{}pos}.
\end{equation}
%
The astrometric cross-match Bayes factor, $B_{\rm{}pos}$, has been derived in \cite{BL15-ARSA} (see Eq.~(17) there, and Eq.~(19) for the tangent plane Gaussian limit that holds for high-precision astrometry).
That work also discusses generalizations that account for proper motion and other complications.

@@

\noindent
We here describe simulations that demonstrate the detection capability of our multi-epoch approach in a setting with known ground truth.
The simulation parameters were chosen to produce data similar to that provided by modern large-scale optical surveys.

%--------------------------------------------------------------------------------
\subsection{Simulated Galaxies}
\noindent
We assume that galaxies are brighter than 28 magnitudes and that the 5$\sigma$ detection limit is 24 magnitudes, corresponding roughly to parameters of LSST photometry.
%
Panel~(b) of Figure~\ref{fig:frac} shows the noise peak detection probability as a function of magnitude based on these parameters, in contrast to the dimensionless presentation in panel~(a).
%
To compute the marginal likelihood for the source-present hypothesis, we must specify a prior for the source flux, $\flux$.
Here we use a standard faint-galaxy number counts law,
with the number counts following the empirical formula of 
\mbox{$dN\!\propto\!10^{0.4m}dm$}; see \cite{MT00-NumCounts}. 
That approximately translates to the properly normalized population distribution of
%
\begin{equation} 
\label{eq:ref3}
\pi(f) = \left\{\begin{array}{l l}
           f_L/f^2  & \quad \mbox{if\ $f > f_L$},\\
           0 & \quad \mbox{otherwise},
           \\ \end{array} \right.
\end{equation}
%
where $f_L$ is the limiting flux that corresponds to the previously defined magnitude limit.

%\subsection{The Mock Catalogs}\label{sec:mock} \noindent
We generate sets of random detections for 20,000 galaxies with true fluxes between 28 and 23 magnitudes by simply drawing $\fest_\eind$ values from a Gaussian centered on the actual fluxes.
%
We also generate 2,000 ghost detection $\fest_\eind$ values from $\npd(\fest)$ by inverting its cumulative distribution (computed numerically on a grid).
%
The number of exposures is set to the previously used $k=9$ with a single-epoch flux threshold of just 1$\sigma$, deep in the noise.
In observations with our specified parameters, the number of ghost detections will greatly outnumber the galaxy detections with this low threshold.
The numbers of galaxies and ghosts were chosen to enable display of the distributions of Bayes factors for the two classes of detections (noise and true).

@@

This computation demonstrates that flux and astrometric catalog data, combined across epochs, can strongly distinguish real objects from spurious detections.
We have not addressed what threshold Bayes factor to use for producing a multi-epoch catalog, or, in Bayesian terms, how to convert Bayes factors into posterior probabilities for candidate objects.
As noted above, when the object population density (on the sky and in flux) is known a priori, the calculation is straightforward (e.g., the prior odds will be proportional to the ratio of true object and noise peak sky densities).
When these quantities are unknown, a possibly complicated hierarchical Bayesian calculation can jointly estimate them and the object properties.
When many objects are detected, an approximate approach, plugging in empirical estimates of the densities based on the data, is likely to suffice, as described in \cite{B12-SCMAXMatch,L12-HierXMatch}.

%****************************************
\color{red}
\begin{figure}
\epsscale{1.2}
\plotone{fig/f7u-U.png}
\caption{The (red) points indicate the weight of evidence for simulated galaxies as a function of their true brightness and the same for generated random noise peaks.
The Bayes factor is high for the bright galaxies but its logarithm is close to zero for the faint ones.
The noise peaks (left of double lines) seem to separate slightly from the faintest sources (just right of the double dashed vertical lines) because the fluxes in noise peaks tend to be higher, which better mimics sources between 26 and 25 magnitudes.}
\label{fig:bf}
\end{figure}

@@

%================================================================================
\section{Summary}
\label{sec:sum}

\noindent
This paper presents an exploratory study of a new, incremental approach to the analysis of multi-epoch survey data, based on fusion of single-epoch catalogs produced using a source detection algorithm with a modest or low threshold.
Although the single-epoch catalogs will include many noise sources (they may even be dominated by them), we show that probabilistic fusion of the single-epoch data can produce interim or final multi-epoch catalogs with properties similar to those expected from catalogs based on image stacking.
The approach is essentially a generalization of cross-matching, where object detection corresponds to identifying a set of candidate sources that match in both flux and direction across epochs.
Using a probabilistic approach directly provides the required quantities, enabling fusion of information both across epochs, and between flux and direction, by straightforward multiplication of the relevant probabilities.
The final quantification of strength of evidence is via marginal likelihoods and Bayes factors; these can be used for final thresholding, or for producing posterior probabilities for source detections when population properties such as sky densities are known or can be accurately estimated (perhaps as part of the catalog analysis).

The Bayes factor compares predictions of the observed data based on true-object and noise-peak hypotheses for the data, and thus requires knowledge of the distribution of noise peaks.
We derive the spatial properties of noise peaks that commonly appear in catalogs.
The flux-dependent surface density of ghosts is asymmetric in flux, skewed toward positive flux values.
It can be accurately approximated by a shifted Gaussian for most practical purposes.

Based on the Bayes factor, sources with single-epoch measured fluxes over a \mbox{3$\sigma$} threshold \mbox{($m\!\simeq\!24.55$ for an LSST-like survey)} start to separate out from the noise peaks when data are combined across just a few epochs.
The evidence for a source becomes very strong once the single-epoch fluxes exceed \mbox{5$\sigma$} (24 mag).
%
When considering only the flux measurements, the faintest sources are hard to distinguish from the noise peaks with measurements at just a few epochs; but astrometric data (celestial coordinate estimates) greatly help to separate genuine and spurious detections.

In general, the specificity and the selectivity of the proposed discriminator depends on a number of parameters, most of which we discuss as part of the simulated case study.
The pixel size, the single-epoch detection threshold, the number of exposures, and the point-spread function all affect the frequency of noise peaks.
The population distribution of source properties also directly impacts detectability of faint sources.
A hierarchical generalization of the approach could learn important features of the population distribution as part of the analysis.

\enote{Not clear what this is getting at, in the previous revision:  ``as well as the statistics of the real sources based on which one can come up with reliable thresholds for progressively weeding out noise from the catalogs throughout the lifetime of a survey.''}

We have treated only the case of detection of constant-flux sources.
Detecting variable and transient sources can be accommodated by introducing one or more time series models into the flux matching part of the algorithm.
Models that accurately describe particular classes of sources will produce optimal catalogs, but flexible models---perhaps simple stochastic processes, or even histogram or other partion-based models, with appropriate priors on variability---may suffice for producing general-purpose multi-epoch catalogs for studying variable sources.
This is a potentially complicated generalization of our framework that we plan to explore in future work.
Of course, variable source detection using image stacks is also an open research problem; our framework provides an alternative avenue to address it.

Our exploratory study made simplifying assumptions.
A strong assumption was that source candidates are isolated enough that the image space can be partitioned into patches that have at most one candidate source.
When source candidates are close to each other, the matching across epochs must account for multiple possible source association hypotheses.
Similar complications appear when considering classes of objects that may be comprised of multiple sources per epoch, e.g., radio galaxies.
In other work, we have developed techniques for directional cross-matching in contexts with multiple candidate associations, and with complex object structure and object motion \citep{BL15-ARSA}.
These methods can be extended to include flux matching criteria to generalize the multi-epoch detection framework described here.

The strategy we have described is quite different from conventional approaches to producing survey catalogs.
Implementing it will raise new processing and database management challenges; users of the resulting catalogs will need to think about catalogs in a different way.
In particular, a low-threshold single-epoch catalog will contain many spurious sources; with a low enough threshold, the spurious sources will greatly outnumber real sources.
However, evidence mounts quickly as catalogs are merged.
If interim catalogs are produced consecutively, cumulative culling of early single-epoch catalogs could reduce the storage burden for catalogs subsequent to the first catalog.
Such issues, and the generalizations described above, will be topics for future study.

@@

\bibitem[Budav{\'a}ri(2012)]{B12-SCMAXMatch} Loredo, T.~J.\ 2012, {Statistical {Challenges} in {Modern} {Astronomy} {V}}, 303--308

\bibitem[Budav{\'a}ri \& Loredo(2015)]{BL15-ARSA} Budav{\'a}ri, T., \& Loredo, T.~J.\ 2015, Annual Review of Statistics and Its Application, 2, 113--139

%\bibitem[Gregory \& Loredo(1992)]{gregory} Gregory, P.~C., \& Loredo, T.~J.\ 1992, \apj, 398, 146

@@

\bibitem[Loredo(2012)]{L12-HierXMatch} Loredo, T.~J.\ 2012, {Statistical {Challenges} in {Modern} {Astronomy} {V}}, 303--308

@@

\appendix
